Enabling GPU Virtualization in Cloud Environments

### 2015-10-26 paper.tex
hasta el 5.1 re-revisado

\section{\uppercase{GPGPU share service with AWS}}
\label{sec:workingaws}
In this section, we briefly expose the current user-level features 
provided by AWS and then analyze this service  
from the point of view of the GPU usage. 
Next, we describe the scenarios and experiments where virtualized GPUs are utilized and show a number of first results.

\subsection{Current Features}
This subsection identifies the most common features checked during 
the VM's creation step.

@@

An instance is a pre-configured VM focused on an 
specific target. Among the large list of instances offered by AWS, 
we can find cases for general purpose (T2, M4 and M3); computer science optimized 
(C4 and C3); memory optimized (R3); storage optimized (I2 and D2) and 
GPU capable (G2). Each type of instance has its own purpose cost (price). 
Moreover, each instance offers a different number of CPUs as well as network 
interconnection which can be: low, medium, high or 10Gb.

For our study, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
The instances available in that zone presented the features shown in the Table~\ref{table:awsInstances}.

@@

For the following experiments, we select C3 family instances, which are not equipped with GPUs, as clients; whereas instances of the G2 family will act as servers.

\subsubsection{Networking}
Table~\ref{table:awsInstances} shows that each instance integrates a different 
network. Thus, taking into account that bandwidth 
is critical when GPU virtualization is applied, a simple test to verify the 
real network bandwidth is needed.

@@

To evaluate the actual bandwidth, we executed 
the IPERF\footnote{\url{http://iperf.fr/}} tool between the instances explained in 
the Table~\ref{table:awsInstances} with the results shown in the Table~\ref{table:iperf}.
From this experiment, we can assume that the nomenclature ``High'' corresponds to a 1~Gb interconnection network and the 10~Gb has 
a real bandwidth of 7.5~Gb/s.
Moreover, it looks like the instances equipped with a ``High'' interconnection network
are constrained by software to 1~Gb/s since the theoretical and real bandwidth 
match perfectly. The real gap between sustantiable and theoretical bandwidth can be observed with 
the 10~Gb interconnection, which reaches up to 7.5~Gb/s.

\subsubsection{GPUs}
An instance relies on a VM that runs on a real node with its own virtualized components. 
Therefore it is possible that AWS uses a virtualization framework to offer GPU services to all the instances.
Although, the {\tt nvidia-smi} command indicates that the GPUs installed are NVIDIA GRID K520, 
we need to verify that they are non-virtualized devices. For this proposal, we execute the NVIDIA SDK test {\tt bandwidthtest}. 
As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwith, revealing that the accelerator is an ad-hoc GPU. 
% \begin{table}[htb]
% \renewcommand{\arraystretch}{1.3}
% \caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
% \label{table:bwt}
% \tabcolsep=0.09cm
% \begin{center}\begin{tabular}{cccc}
% Name &  Data Movement & Network & Bandwidth \\ \hline \hline
% g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
% g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
% g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
% g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
% \end{tabular}\end{center}\end{table}

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory in local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{ccc}
Name &  Data Movement & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

\subsection{Testbed Scenarios}
In this section, we introduce the different client-server configurations. 
All scenarios are based on the maximum number of instances 
that a user can freely select without submitting a formal request. 
In particular, the maximum number for ``g2.2xlarge'' is 5 and for ``g2.8xlarge'' it is 2.

All the instances deployed operate the RHEL 7.1 64-bit OS and version 6.5 of CUDA. 
We design three configuration scenarios for our tests:

\begin{itemize}
\item Scenario A (Figure~\ref{subfig:aws1}) shows the most common 
 hardware configuration in clusters, with each node  
populated with one GPU. Here, a simple node can access to 5 GPUs 
using the ``High'' network.

\item Scenario B (Figure~\ref{subfig:aws2}) is composed by 2 server nodes equipped 
with 4 GPUs each one, and a GPU-less client. This scenario includes 
a 10Gb network, and the client can execute the application using up to 8 GPUs.

@@

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32 MB using pageable memory using {rCUDA}.}
\label{table:bwtrcuda}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Scenario &  Data Movement & Network & Bandwidth \\ \hline \hline
A & Host-to-Device & High& 127 MB/s \\ \hline
A & Device-to-Host & High& 126 MB/s\\ \hline
B & Host-to-Device & 10 Gb& 858 MB/s\\ \hline
B & Device-to-Host & 10 Gb& 843 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

>>>>

\section{\uppercase{GPGPU share service with AWS}}
\label{sec:workingaws}
In this section, we briefly expose the current user-level features 
provided by AWS, and then analyze this service  
from the point of view of GPU usage. 
Next, we describe the scenarios and experiments where virtualized GPUs are utilized and show a number of initial results.

\subsection{Current Features}
This subsection identifies several common features checked during 
the VM's creation step.

@@

An instance is a pre-configured VM focused on an 
specific target. Among the large list of instances offered by AWS, 
we can find specialized versions for general-purpose (T2, M4 and M3); computer science (C4 and C3); memory (R3); storage (I2 and D2) and 
GPU capable (G2). Each type of instance has its own purpose and cost (price). 
Moreover, each type offers a different number of CPUs as well as network 
interconnection, which can be: low, medium, high or 10Gb.

For our study, we worked in the AWS availability zone US EAST (N. VIRGINIA). 
The instances available in that case present the features reported in Table~\ref{table:awsInstances}.

@@

For the following experiments, we select C3 family instances, which are not equipped with GPUs, as clients; whereas instances of the G2 family will act as GPU-enabled servers.

\subsubsection{Networking}
Table~\ref{table:awsInstances} shows that each instance integrates a different 
network. As the bandwidth 
is critical when GPU virtualization is applied, we first perform a simple test to verify the 
real network bandwidth.

@@

To evaluate the actual bandwidth, we executed 
the IPERF\footnote{\url{http://iperf.fr/}} tool between the instances described in 
Table~\ref{table:awsInstances}, with the results shown in Table~\ref{table:iperf}.
From this experiment, we can derive that network ``High'' corresponds to a 1~Gb interconnect while ``10~Gb'' has 
a real bandwidth of 7.5~Gb/s.
Moreover, it seems that the bandwidth of the instances equipped with a ``High'' interconnection network
is constrained by software to 1~Gb/s since the theoretical and real bandwidth 
match perfectly. The real gap between sustained and theoretical bandwidth can be observed with 
the 10~Gb interconnection, which reaches up to 7.5~Gb/s.

\subsubsection{GPUs}
An instance relies on a VM that runs on a real node with its own virtualized components. 
Therefore AWS can leverage a virtualization framework to offer GPU services to all the instances.
Although the {\tt nvidia-smi} command indicates that the GPUs installed are NVIDIA GRID K520, 
we need to verify that these are non-virtualized devices. For this purpose, we execute the NVIDIA SDK {\tt bandwidthtest}. 
As shown in the Table~\ref{table:bwt}, the bandwidth achieved in this test is higher than the network bandwith, revealing that the accelerator is an actual GPU. 
% \begin{table}[htb]
% \renewcommand{\arraystretch}{1.3}
% \caption{NVIDIA SDK {\tt bandwidthtest} execution transferring 32MB using pageable memory using local GPU.}
% \label{table:bwt}
% \tabcolsep=0.09cm
% \begin{center}\begin{tabular}{cccc}
% Name &  Data Movement & Network & Bandwidth \\ \hline \hline
% g2.2xlarge & Host to Device & High& 3,004 MB/s \\ \hline
% g2.2xlarge & Device to Host & High& 2,809 MB/s\\ \hline
% g2.8xlarge & Host to Device & 10 Gb& 2,814 MB/s\\ \hline
% g2.8xlarge & Device to Host & 10 Gb& 3,182 MB/s\\ \hline
% \end{tabular}\end{center}\end{table}

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Execution of NVIDIA SDK {\tt bandwidthtest} to transferring 32MB using pageable memory in a local GPU.}
\label{table:bwt}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{ccc}
Name &  Data Movement & Bandwidth \\ \hline \hline
g2.2xlarge & Host to Device & 3,004 MB/s \\ \hline
g2.2xlarge & Device to Host & 2,809 MB/s\\ \hline
g2.8xlarge & Host to Device & 2,814 MB/s\\ \hline
g2.8xlarge & Device to Host & 3,182 MB/s\\ \hline
\end{tabular}\end{center}\end{table}

\subsection{Testbed Scenarios}
In this section, we introduce the different client-server configurations. 
All scenarios are based on the maximum number of instances 
that a user can freely select without submitting a formal request. 
In particular, the maximum number for ``g2.2xlarge'' is 5; for ``g2.8xlarge'' it is 2.

All instances operate the RHEL 7.1 64-bit OS and version 6.5 of CUDA. 
We design three configuration scenarios for our tests:

\begin{itemize}
\item Scenario A (Figure~\ref{subfig:aws1}) shows a common 
configuration in GPU-accelerated clusters, with each node  
populated with a single GPU. Here, a node can access 5 GPUs 
using the ``High'' network.

\item Scenario B (Figure~\ref{subfig:aws2}) is composed of 2 server nodes, equipped 
with 4 GPUs each, and a GPU-less client. This scenario includes 
a 10Gb network, and the client can execute the application using up to 8 GPUs.

@@

\begin{table}[htb]
\renewcommand{\arraystretch}{1.3}
\caption{Execution of NVIDIA SDK {\tt bandwidthtest} to transferring 32MB using pageable memory in a remote GPU using {rCUDA}.}
\label{table:bwtrcuda}
\tabcolsep=0.09cm
\begin{center}\begin{tabular}{cccc}
Scenario &  Data Movement & Network & Bandwidth \\ \hline \hline
A & Host-to-Device & High& 127 MB/s \\ \hline
A & Device-to-Host & High& 126 MB/s\\ \hline
B & Host-to-Device & 10 Gb& 858 MB/s\\ \hline
B & Device-to-Host & 10 Gb& 843 MB/s\\ \hline
\end{tabular}\end{center}\end{table}
